\bf An Examination of Sport Climbing Competition Format and Scoring System

### 2021-08-23 manuscript/Rmd/draft.Rmd
Start PCA

@@

@@

>>>>

Paragraph on rank swapping

Paragraph on PCA

@@

(Need a more detailed description)

@@

## Rank Swapping Analysis

## Principal Component Analysis

Olympic results

```{r}
# Women's qual
wq <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2020_olympics/wq.csv")
wq_cleaned <- wq %>%
  mutate(
    a = as.numeric(str_remove_all(a, "[A-z]")),
    b = as.numeric(str_remove_all(b, "[A-z]")),
    speed_best = ifelse(is.na(a), b,
                        ifelse(is.na(b), a,
                               ifelse(a < b, a, b))),
    bould_tops = as.numeric(str_sub(results, 1, 1)),
    lead_hr = ifelse(
      str_detect(hr, "\\+"),
      as.numeric(str_sub(hr, 1, nchar(hr) - 1)) + 0.5,
      as.numeric(hr)
    )
  ) %>%
  select(climber, speed_best, bould_tops, lead_hr) %>%
  column_to_rownames(var = "climber")

# Men's qual
mq <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2020_olympics/mq.csv")
mq_cleaned <- mq %>%
  mutate(
    a = as.numeric(str_remove_all(a, "[A-z]")),
    b = as.numeric(str_remove_all(b, "[A-z]")),
    speed_best = ifelse(is.na(a), b,
                        ifelse(is.na(b), a,
                               ifelse(a < b, a, b))),
    bould_tops = as.numeric(str_sub(results, 1, 1)),
    lead_hr = ifelse(
      str_detect(hr, "\\+"),
      as.numeric(str_sub(hr, 1, nchar(hr) - 1)) + 0.5,
      as.numeric(hr)
    )
  ) %>%
  select(climber, speed_best, bould_tops, lead_hr) %>%
  column_to_rownames(var = "climber")
```

```{r}
library(broom)
```

```{r, fig.cap= "Women's Qualification PCA", fig.width = 6, fig.height = 5}
# pca fit
wq_pca <- wq_cleaned %>%
  prcomp(scale = TRUE)

# plot
library(ggfortify)
wq_pca %>% 
  autoplot(label = TRUE, 
           label.size = 3,
           label.repel = TRUE,
           loadings = TRUE,
           loadings.label = TRUE, 
           loadings.label.size = 5)
```

```{r, fig.cap= "Men's Qualification PCA", fig.width = 6, fig.height = 5}
# pca fit
mq_pca <- mq_cleaned %>%
  prcomp(scale = TRUE)

# plot
library(ggfortify)
mq_pca %>% 
  autoplot(label = TRUE, 
           label.size = 3,
           label.repel = TRUE,
           loadings = TRUE,
           loadings.label = TRUE, 
           loadings.label.size = 5)
```

### 2021-08-23 manuscript/Rmd/draft.tex
Start PCA

@@

{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-16-1}

@@

\caption{Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications}\label{fig:unnamed-chunk-16}
\end{figure}

@@

{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-19-1}

@@

\caption{This figure illustrates the changes to the final rankings of the 2018 Youth Olympics Men's and Women's Finals when we leave out one climber. For each gender competition, each panel represents the rank of the drop-out athlete, with 0 being the original final results. Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar}\label{fig:unnamed-chunk-19}
\end{figure}

@@

\caption{\label{tab:unnamed-chunk-21}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.5731 & 0.6760 & 0.6126 & 1.0000 & 0.5474 & 0.6878 & 0.7053\\
Speed & 0.5731 & 1.0000 & 0.3022 & 0.2174 & 0.5474 & 1.0000 & 0.2540 & 0.2947\\
Bouldering & 0.6760 & 0.3022 & 1.0000 & 0.6203 & 0.6878 & 0.2540 & 1.0000 & 0.6878\\
Lead & 0.6126 & 0.2174 & 0.6203 & 1.0000 & 0.7053 & 0.2947 & 0.6878 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-22}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.2632 & 0.3684 & 0.3474 & 1.0000 & 0.2316 & 0.5013 & 0.2526\\
Speed & 0.2632 & 1.0000 & -0.0947 & -0.3263 & 0.2316 & 1.0000 & 0.0686 & -0.4526\\
Bouldering & 0.3684 & -0.0947 & 1.0000 & 0.2000 & 0.5013 & 0.0686 & 1.0000 & 0.2902\\
Lead & 0.3474 & -0.3263 & 0.2000 & 1.0000 & 0.2526 & -0.4526 & 0.2902 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-23}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.0947 & 0.3810 & 0.2632 & 1.0000 & 0.1847 & 0.5358 & 0.5368\\
Speed & 0.0947 & 1.0000 & -0.1587 & -0.4316 & 0.1847 & 1.0000 & -0.1915 & -0.0897\\
Bouldering & 0.3810 & -0.1587 & 1.0000 & 0.2751 & 0.5358 & -0.1915 & 1.0000 & 0.5040\\
Lead & 0.2632 & -0.4316 & 0.2751 & 1.0000 & 0.5368 & -0.0897 & 0.5040 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-24}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.5906 & 0.5556 & 0.6491 & 1.0000 & 0.5789 & 0.6140 & 0.5906\\
Speed & 0.5906 & 1.0000 & 0.2164 & 0.4035 & 0.5789 & 1.0000 & 0.2632 & 0.3099\\
Bouldering & 0.5556 & 0.2164 & 1.0000 & 0.2749 & 0.6140 & 0.2632 & 1.0000 & 0.3918\\
Lead & 0.6491 & 0.4035 & 0.2749 & 1.0000 & 0.5906 & 0.3099 & 0.3918 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

>>>>

Olympic Finalist Bassa Mawem pulled out of the sport climbing finals due
to an injury in his left arm. Reuters reported that Megos, originally
9th in the total qualification rankings would not be replacing him. We
suspect that this might be part of the unclear format of the Olympic
Sport Climbing format. However, according to the climber's rankings,
Megos technically should not have been next in line to fill in the
missing finalist spot. Upon removing B. Mawem from the list and
reranking the remaining competitors, we find that Chon would have moved
into 8th place with a 640 combined score and Megos would have remained
in 9th with 648. B. Mawem originally ranked first in the speed climbing,
and Megos did comparatively poorly with 19th place. Both performed
better than Mawem in bouldering and lead, leaving those rankings
unchanged.

Prior, we considered a change in finalists if a non-finalist was
removed. It is clear now though that the more interesting scenario is
would would replace an original finalist who is removed from the
rankings. Should it be the original 9th place competitor, or does it
make more sense with such a scoring system to rerank? Surely Chon should
be rewarded for his superior performance in speed compared to Megos with
the chance to win a medal?

The next set of simulations will evaluate the probability that a removed
finalist is replaced with someone other than the original 9th place
competitor.

\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-14-1} \end{center}

@@

{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-17-1}

@@

\caption{Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications}\label{fig:unnamed-chunk-17}
\end{figure}

@@

{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-20-1}

@@

\caption{This figure illustrates the changes to the final rankings of the 2018 Youth Olympics Men's and Women's Finals when we leave out one climber. For each gender competition, each panel represents the rank of the drop-out athlete, with 0 being the original final results. Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar}\label{fig:unnamed-chunk-20}
\end{figure}

\hypertarget{rank-swapping-analysis}{%
\subsection{Rank Swapping Analysis}\label{rank-swapping-analysis}}

\hypertarget{principal-component-analysis}{%
\subsection{Principal Component
Analysis}\label{principal-component-analysis}}

Olympic results

{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-23-1}

\caption{Women's Qualification PCA}\label{fig:unnamed-chunk-23}
\end{figure}

{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-24-1}

\caption{Men's Qualification PCA}\label{fig:unnamed-chunk-24}
\end{figure}

@@

\caption{\label{tab:unnamed-chunk-26}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.5731 & 0.6760 & 0.6126 & 1.0000 & 0.5474 & 0.6878 & 0.7053\\
Speed & 0.5731 & 1.0000 & 0.3022 & 0.2174 & 0.5474 & 1.0000 & 0.2540 & 0.2947\\
Bouldering & 0.6760 & 0.3022 & 1.0000 & 0.6203 & 0.6878 & 0.2540 & 1.0000 & 0.6878\\
Lead & 0.6126 & 0.2174 & 0.6203 & 1.0000 & 0.7053 & 0.2947 & 0.6878 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-27}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.2632 & 0.3684 & 0.3474 & 1.0000 & 0.2316 & 0.5013 & 0.2526\\
Speed & 0.2632 & 1.0000 & -0.0947 & -0.3263 & 0.2316 & 1.0000 & 0.0686 & -0.4526\\
Bouldering & 0.3684 & -0.0947 & 1.0000 & 0.2000 & 0.5013 & 0.0686 & 1.0000 & 0.2902\\
Lead & 0.3474 & -0.3263 & 0.2000 & 1.0000 & 0.2526 & -0.4526 & 0.2902 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-28}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.0947 & 0.3810 & 0.2632 & 1.0000 & 0.1847 & 0.5358 & 0.5368\\
Speed & 0.0947 & 1.0000 & -0.1587 & -0.4316 & 0.1847 & 1.0000 & -0.1915 & -0.0897\\
Bouldering & 0.3810 & -0.1587 & 1.0000 & 0.2751 & 0.5358 & -0.1915 & 1.0000 & 0.5040\\
Lead & 0.2632 & -0.4316 & 0.2751 & 1.0000 & 0.5368 & -0.0897 & 0.5040 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-29}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.5906 & 0.5556 & 0.6491 & 1.0000 & 0.5789 & 0.6140 & 0.5906\\
Speed & 0.5906 & 1.0000 & 0.2164 & 0.4035 & 0.5789 & 1.0000 & 0.2632 & 0.3099\\
Bouldering & 0.5556 & 0.2164 & 1.0000 & 0.2749 & 0.6140 & 0.2632 & 1.0000 & 0.3918\\
Lead & 0.6491 & 0.4035 & 0.2749 & 1.0000 & 0.5906 & 0.3099 & 0.3918 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}
